% Preamble
\documentclass[../main.tex]{subfiles}

% Document
\begin{document}
\chapter{Introduction}\label{ch:introduction}

In this work it is intended to address the multi-armed bandit problem and its
extension to contextual multi-armed bandits. This problem is one of the most common
examples within reinforcement learning and has numerous applications in the
real world. Of these applications, the most widespread are: the design of
recommender systems (for streaming platforms, e-commerce, or social
networks), clinical trials, dynamic routing, and the design of A/B tests.  

In the case of recommender systems, the boom in the number of services that
nowadays are carried out by most people through the internet has made these systems
take on enormous importance when it comes to maximizing profits for multiple companies
and, therefore, the application of algorithms for the resolution of the multi-armed
bandit problem and reinforcement learning in general has gained a greater degree of
relevance recently.  

The majority of these services offered through the internet generate an enormous
quantity of data about how users interact with a web page, which causes this type of
application of the multi-armed bandit problem to be carried out in Big Data
environments, bringing certain challenges, although also certain advantages, since
reinforcement learning algorithms never stop learning, and as more data are generated
(as long as these data are of good quality) these systems will continue improving more
and more every day.  

With respect to this work, it is intended to carry out a theoretical introduction to
the multi-armed bandit problem and its extension to the contextual multi-armed bandit,
and also to the main algorithms used to solve these problems, additionally showing how
these algorithms would be used for the recommender system of an e-commerce
platform.  

The document will be structured in the following way:
\begin{itemize}
\item In a first section, the motivation and the objectives of the work will be
presented.
\item Next, a review of the literature will be made, in which the entire theoretical
context of these problems and the main algorithms used to solve them will be
explained.
\item Afterward, there will be a section devoted to the methodology followed for the
implementation of the practical part of the work, in which the algorithms presented in
the state of the art will be applied to the recommender system on an e-commerce
website.
\item After the methodology, there will be a section of results and discussion, in
which the results obtained from the application of these algorithms to the
recommender system will be explained.
\item And to conclude, there will be a conclusions section in which the general
conclusions of the work will be presented.
\end{itemize}

\end{document}